Improving Item-Item Similarity Estimation during Collaborative Filtering

### 2014-03-26 paper/paper.tex
Make minor tweaks.

\section*{Abstract}
\emph{The problem of inaccurate item-item similarity estimates due to a lack of
data is explored within the context of collaborative filtering. An off-line
experiment that evaluates the performance of similarity estimators is described.
Two novel item-item similarity estimators are described and analyzed. The first
is based on a model of noisy similarity score generation and proves effective at
estimating both future Pearson correlations and cosine similarities. The second
models noisy ratings as generated by similarity score dependent distributions.
Using the notion of user predictivity, accurate estimates of future Pearson
correlation are made.}

\section*{Introduction}
Determining the similarity of two users or items is valuable for a number of
data-mining applications including recommender systems, product assortment and
differentiation, as well as branding. Collaborative filtering is often used to
exploit user-item rating data to make such estimates. This is typically done by
computing the Pearson correlation or cosine similarity of vector representations
of the users or items \cite{Su2009}. Given sufficient user-item rating data,
this approach can be very effective. However, when the rating data is sparse for
one or both users/items, these metrics can yield dramatic discrepancies compared
to values computed with sufficient data at a later point in time. Moreover,
these early, low-confidence measurements can be indistinguishable from later,
high-confidence measurements.  In this report, we describe two novel techniques
for estimating the true similarity of two users or items given limited common
rating data by incorporating confidence into the measurement.

@@

\section*{Similarity}
\subsection*{True Item-Item Similarity}

@@

In the second model, we imagine an unlimited supply of users, each of whom
chooses to rate a subset of items. The lack of a rating is now meaningful since
it indicates a lack of interest of a particular user for a particular item. This
is captured by the model through a default rating of zero\footnote{We address
the issue of rating bias and the precise meaning of a zero valued ratings later
in this report.}. At any point in time, we may estimate the true similarity in
this model of {\em preferential rating}, by computing the cosine similarity of
the two item vectors. Only in the limit of inifinite time does this estimate
converge to the true similarity, which we denote $TrueSimPref$.  That is,
\begin{align}
TrueSimPref(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}},
\end{align}
where $U_A$ is the set of users who have rated $A$, $U_B$ is the set of users
who have rated $B$, and $U_{AB}$ is the set of users who have rated both $A$ and
$B$.

\subsection*{Estimating Item-Item Similarity}

Of course, most systems will never acquire ratings from all users for all items.
Nor are we able to work in the limit of infinite time. We must therefore make
estimates of the true item-item similarity using only partial information. In
the case of $TrueSimPref$, the cosine similarity computed at any point in time
serves as a simple and obvious estimate. In using this estimate, we are treating
all missing ratings as intentionally omitted. When estimating $TrueSimRand$, the
choice of how to handle missing ratings is perhaps less obvious. Following the
convention of Breese et al. \cite{Breese1998}, we choose to use the Pearson
correlation\footnote{The Pearson correlation is traditionally characterized by
subtracting an expectation value from each dimension. This is only approximately
true in our case. We discuss the how we subtract rating biases later in this
report.} for this estimate, ignoring all ratings by users who have not rated
both items. To summarize, our naive estimates are
\begin{align}
PearSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{AB}} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_{AB}} r_{u,B}^2}}
\stackrel{?}{\approx} TrueSimRand(A, B)
\end{align}

@@

\section*{Problem}
Both similarity estimates primarily leverage information from common users, that
is users who have rated both items. Because of this, they perform well with a
large number of common users, but give unreliable, and dramatically varying,
results when the items have few common users. In an extreme case, if only one
user has rated item $A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is
this unlikely to be an accurate assessment of the true similarity of $A$ and
$B$, it also gives the most extreme results possible, without giving any
indication that this is a low-confidence calculation.

@@

\subsection*{Data}
In the following section, we will describe an off-line experiment used to
evaluate our similarity estimators. But first, we provide some details about the
data used for this work. The data, originally from Amazon.com, was acquired from
the Stanford Network Analysis Project\footnote{http://snap.stanford.edu}. It
consists of ~5.6M ratings by ~1.2M users of ~600k music items (e.g. albums and
songs). Each rating is from 1 to 5 stars and is associated with a particular
user and item at a particular time.  For practical reasons, we focus our work on
two subsets of this data. The first consists of the 13,752 item pairs having at
least 40 common users (i.e. users that have provided a rating for the item). The
second consists of the 1,312 item pairs having at least 80 common reviewers. As
discussed below, we must confine ourselves to item pairs that have many common
reviewers in order to determine a good approximation of the their true
similarity.

@@

To obtain good $GoldSim$ values, we work only with the item pairs that have at 
least either $N=40$ and $80$ common users. These choices are a compromise 
between having enough data for statistically significant results and having 
a sufficiently large $N$ that $GoldSim \approx TrueSim$.

In order to test predicted $TrueSim$, we construct an off-line experiment to
simulate observing sparse amounts of rating data. Our experiment reenacts  
the passing of time. Starting with all ratings withheld, we apply ratings to 
items in the order they actually occured. The similarity of item pairs is then
estimated each time the number of common users $n$ is incremented. In this way,
we can quanitatively compare various estimators by calculating their mean
squared error when approximating $GoldSim$ across pairs of items. Since 
different estimators may perform better for different degrees of sparsity, 
we compute the mean squared error for $n=1,...,N$ separately.

@@

In our first approach, we model the probability distributions of true and 
observed similarity. The goal is to use the observed similarity score directly 
to approximate the true similarity. In this way, we can consider an abstact 
$TrueSim$ which ignores the details of $TrueSimRand$ or $TrueSimPref$. 
Similarly, we consider $Sim_n$ to be a generic observed similarity score when
the two items have $n$ common reviewers. Now, let $Y$ be a random variable 
representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
random variable representing the $Sim_{n}$ of a pair of items when $n$ common 
users are observed.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear predictors. For each number of
reviewers in common, $n$, the graphs represent the average squared error of
predicted $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the
technique of guessing the average $GoldSim$ is included. For these experiments,
we use $Sim$ = $PearSim$ and $N=80$ for determining $GoldSim$.}
    \label{fig:Linear}
\end{figure}

@@

Let $Y$ be a random variable representing the $TrueSimRand$ of a randomly chosen
pair of items. As before, we model $Y$ as a Normal distribution
\begin{align}
Y \sim N(\mu_s, \sigma_s^2),
\end{align}
where $\mu_s$ and $\sigma_s^2$ are the average and variance of all true
similarity scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be a
random two element vector representing the ratings given to items $A$ and $B$.
We model the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
distribution
\begin{align}
(\vec{R} | Y=y) \sim N(\vec{\mu_r}, \mathbf{\Sigma_y}),
\end{align}
where $\vec{\mu_r} = (\mu_r~~\mu_r)$ are the average ratings and
$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix.  Since
$y$ is the $TrueSimRand$ score, it can be thought of as the Pearson correlation
over the distribution of ratings $R_A | Y$ and $R_B | Y$. Assuming that the
variance of individual ratings $\sigma_r^2$ is independent of $y$, we have 
\begin{align}
y = \frac{\Cov{\left[R_A | Y=y , R_B | Y=y\right]}}{\sigma_r^2},
\end{align}
It therefore follows that if the ratings are indeed normally distributed 
then the covariance matrix is related to the true similarity $y$ by
\begin{align}
\mathbf{\Sigma_y} = \sigma_r^2 \cdot
\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right].
\end{align}

@@

As already mentioned, the estimator $PearSim$ is unable to say much about the
similarity of two items when only provided ratings from a single common user. In
this case, $PearSim$ yields $\pm1$ (or the indeterminate value $0/0$ in the case
that both ratings are zero) depending on whether the ratings have the same sign
or not. The value of $\hat{y}$ provided by the above model rectifies this by
generating similarity estimates in the range $[-1, +1]$ given a single pair of
ratings. We therefore consider each user who rates a pair of items as implicitly
predicting the similarity of the items. A key discovery of the present work is
that the accuracy of users' predictions in a training set is correlated with the
accuracy in a test set. We exploit this finding by ascribing a quantity we call
{\em user predictivity} to each user defined as
\begin{align}\label{eqn:predictivity}
p_u = e^{-\epsilon_u/\lambda},
\end{align}
where $\lambda$ is a parameter to be learned and
\begin{align}\label{eqn:userError}
\epsilon_u = \frac{1}{\left|I_u\right|}
\sum_{A,B \in I_u}{\left|\hat{y}(r_{u,A},r_{u,B}) - GoldSim(A, B)\right|}
\end{align}
with $I_u$ the set of item pairs rated by user $u$. We are once again using
$GoldSim$ as a proxy for $TrueSim$ (See {\bf Experiment} section) such that
$\epsilon_u$ is the average absolute error of user $u$'s similarity predictions.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{RatingDist_40.png}
    \includegraphics[width=0.7\textwidth]{RatingDist_80.png}
	\caption{Distribution of rating pairs by common reviewers for item pairs.
Top figure shows distribution for item pairs having at least 40 common
reviewers. Bottom figure shows distribution for item pairs having at least 80
common reviewers. Rating1 and Rating2 are chosen in arbitrary order.}
    \label{fig:RatingDist}
\end{figure}

@@

Having chosen values for $\mu_s$, $\sigma_s$, $\mu_r$, and $\sigma_r$ we are now
able to estimate the similarity of item pairs. To do this, we must solve Eq.
(\ref{eqn:estimator}) for $\hat{y}$ given a pair of ratings. This can be easily
done in practice using a root finding algorithm\footnote{To solve Eq.
(\ref{eqn:estimator}) for $\hat{y}$ given ratings $r_A$ and $r_B$, we use the
{\tt brentq} root finding function of the widely available Python module {\tt
scipy.optimize}.}. The implied predictions of individual users can then be
aggregated to form a new estimate of the similarity of items $A$ and $B$ as
\begin{align}\label{eqn:userPredSim}
UserPredSim(A,B) = \frac{\sum_{u \in U_{AB}}{p_u\cdot\hat{y}(r_{u,A}, r_{u,B})}}
{\sum_{u \in U_{AB}}{p_u}},
\end{align}
where, as before, $U_{AB}$ is the set of common users for items $A$ and $B$ and
$p_u$ is the {\em user predictivity} defined in Eq. (\ref{eqn:predictivity}). We
leverage the training set to find $p_u$ for each user $u$. This is done by
recording the implied prediction $\hat{y}(r_{u,A},r_{u,B})$ for each common user
$u$ for each item pair $(A, B)$ in the training set and then computing each
user's average absolute error $\epsilon_u$ as per Eq. (\ref{eqn:userError}).
Finally, we apply supervised learning techniques to optimize the value of the
parameter $\lambda$ that appears in the definition of $p_u$. This is achieved by
minimizing the error in Eq. {\ref{eqn:userPredSim} made against the training
set. We find that $\lambda = 0.01$, which yields a strong preference for high
{\em user predictivity}, works well for the data studied here.

Figure \ref{fig:MSE_SimR} shows the results of our experiment using
$UserPredSim$ to estimate $TrueSimRand$. To be clear, these plots are showing
the mean squared error (MSE) given by
\begin{align}
\mathrm{MSE} = \frac{1}{\left|S\right|}\sum_{(A,B) \in S}
{\left(Estimator(A,B) - GoldSim(A,B)\right)^2},
\end{align}
where $S$ is the test set of item pairs and $Estimator(A, B)$ represents an
arbitrary estimator of the similarity between items $A$ and $B$. In the upper
panel, we show the MSE for the $N \geq 40$ data. The black dashed line
represents the estimates of $PearSim$. For example, with $n = 20$ common users,
the MSE of $PearSim$ is about $0.07$, which corresponds to an average absolute
error of $0.26$ in the similarity score---roughly the magnitude of $\sigma_s/2$
(see Fig. \ref{fig:Histograms}). These data clearly show that the reliability of
$PearSim$ in estimating $TrueSimRand$ deteriorates rapidly as $n$ decreases.
Below about $n = 20$, the horizontal red line, which represents the MSE obtained
from always guessing the average $GoldSimRand$ value from the training set,
outperforms $PearSim$. These two estimators serve as simple benchmarks.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.9\textwidth]{MSE_SimR_40.png}
    \includegraphics[width=0.9\textwidth]{MSE_SimR_80.png}
	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
estimates of {\em TrueSimRand} for item pairs ultimately having at least 40
common reviewers (top) and 80 common reviewers (bottom). The dashed curve
represents {\em SimR} computed using only the ratings from the first $n$
reviewers. The flat red line indicate the MSE that results from always guessing
the average {\em TrueSimR} from the training set. The green curve is {\em
LinRegSim} and the blue curve is {\em UserPredSim}. }
    \label{fig:MSE_SimR}
\end{figure}

@@

The lower panel of Fig. \ref{fig:MSE_SimR} shows the MSE of estimators in the
experiment carried out with item pairs with $N \geq 80$. The curves for
$PearSim$, always guessing the average, and $LinRegSim$ exhibit behavior
analogous to that observed in the $N \geq 40$ experiment, with addition of
increased noise. The performance of $UserPredSim$, meanwhile, is significantly
improved, achieving the MSE that $PearSim$ reaches at $n \approx 70$ by $n
\approx 20$. This result demonstrates the efficacy, at least in some datasets,
of measuring and exploiting {\em user predictivity} when estimating item-item
similarity.

@@

The first is based on a model of noisy score observations. Using a series of
linear regressions, this estimator consistently performs as well or better than
both the Pearson correlation or cosine similarity at estimating future values of
these same functions. The second new estimator is based on a model of noisy
rating pairs generated by an item pair having a true, but unknown, similarity.
This method uses a parametrized surface to efficiently estimate item-item
similarity given only a single pair of ratings. Combining these implied
similarity ``predictions'' with the notion of {\em user predictivity}, we
constructed a novel item-item similarity estimator that performs very well, but
whose effectiveness appears limited to estimations of future Pearson
correlation.

@@

In a different direction, we believe there are a number of possible applications
of the user predictivity score presented in this paper. Based on the results
found here, it is reasonable to believe that adding weights to users by their
predictivity would increase the performance of any user-item rating predictor.
The predictivity could also have applications in the problem of Robust
Collaborative Filtering \cite{Mehta2007}, where rating data is particularly noisy
or riddled with spam.

>>>>

\section*{Abstract}
\emph{The problem of inaccurate item-item similarity estimates due to a lack of
data is explored within the context of collaborative filtering. An off-line
experiment that evaluates the performance of similarity estimators is described.
Two novel item-item similarity estimators are described and analyzed. The first
is based on a model of noisy similarity score generation and proves effective at
estimating both future Pearson correlations and cosine similarities. The second
models noisy ratings as generated by similarity score dependent distributions.
Using the notion of user-predictivity, accurate estimates of future Pearson
correlation are made.}

\section*{Introduction}
Determining the similarity of two users or items is valuable for a number of
data-mining applications including recommender systems as well as product
assortment and differentiation. Collaborative filtering is often used to exploit
user-item rating data to make such estimates. This is typically done by
computing the Pearson correlation or cosine similarity of vector representations
of the users or items \cite{Su2009}. Given sufficient user-item rating data,
this approach can be very effective. However, when the rating data is sparse for
one or both users/items, these metrics can yield dramatic discrepancies compared
to values computed with sufficient data at a later point in time. Moreover,
these early, low-confidence measurements can be indistinguishable from later,
high-confidence measurements.  In this report, we describe two novel techniques
for estimating the true similarity of two users or items given limited common
rating data by incorporating confidence into the measurement.

@@

\section*{True Item-Item Similarity}

@@

In the second model, we imagine an unlimited supply of users, each of whom
chooses to rate a subset of items. The lack of a rating is now meaningful since
it indicates a lack of interest of a particular user for a particular item. This
is captured by the model through a default rating of zero\footnote{We address
the issue of rating bias and the precise meaning of a zero valued ratings later
in this report.}. At any point in time, we may estimate the true similarity in
this model of {\em preferential rating}, by computing the cosine similarity of
the two item vectors. Only in the limit of infinite time does this estimate
converge to the true similarity, which we denote $TrueSimPref$.  That is,
\begin{align}
TrueSimPref(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}},
\end{align}
where $U_A$ is the set of users who have rated $A$, $U_B$ is the set of users
who have rated $B$, and $U_{AB}$ is the set of users who have rated both $A$ and
$B$.

\section*{Estimating Item-Item Similarity}

Of course, most systems will never acquire ratings from all users for all items.
Nor are we able to work in the limit of infinite time. We must therefore make
estimates of the true item-item similarity using only partial information. In
the case of $TrueSimPref$, the cosine similarity computed at any point in time
serves as a simple and obvious estimate. In using this estimate, we are treating
all missing ratings as intentionally omitted. When estimating $TrueSimRand$, the
choice of how to handle missing ratings is perhaps less obvious. Following the
convention of Breese et al. \cite{Breese1998}, we choose to use the Pearson
correlation\footnote{The Pearson correlation is traditionally characterized by
subtracting an expectation value from each dimension. This is only approximately
true in our case. We discuss how we subtract rating biases later in this
report.} for this estimate, ignoring all ratings by users who have not rated
both items. To summarize, our naive estimates are
\begin{align}
PearSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{AB}} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_{AB}} r_{u,B}^2}}
\stackrel{?}{\approx} TrueSimRand(A, B)
\end{align}

@@

\section*{Problem}
Both similarity estimates primarily leverage information from common users, that
is users who have rated both items. Because of this, they perform well with a
large number of common users, but give unreliable, and dramatically varying,
results when the items have few common users. In an extreme case, if only one
user has rated item $A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is
this unlikely to be an accurate assessment of the true similarity of $A$ and
$B$, it also gives the most extreme result possible, without giving any
indication that this is a low-confidence calculation.

@@

\subsection*{Data}
In the following section, we will describe an off-line experiment used to
evaluate our similarity estimators. But first, we provide some details about the
data used for this work. The data, originally from Amazon.com, was acquired from
the Stanford Network Analysis Project\footnote{http://snap.stanford.edu}. It
consists of ~5.6M ratings by ~1.2M users of ~600k music items (e.g. albums and
songs). Each rating is from 1 to 5 stars and is associated with a particular
user and item at a particular time.  For practical reasons, we focus our work on
two subsets of this data. The first consists of the 13,752 item pairs having at
least 40 common users (i.e. users that have provided a rating for both items).
The second consists of the 1,312 item pairs having at least 80 common reviewers.
As discussed below, we must confine ourselves to item pairs that have many
common reviewers in order to determine a good approximation of their true
similarity.

@@

To obtain good $GoldSim$ values, we work only with the item pairs that have at
least either $N=40$ or $80$ common users. These choices are a compromise between
having enough data for statistically significant results and having a
sufficiently large $N$ that $GoldSim \approx TrueSim$.

In order to evaluate estimates of $TrueSim$, we construct an off-line experiment
to simulate observing sparse amounts of rating data. Our experiment reenacts the
passing of time. Starting with all ratings withheld, we apply ratings to items
in the order they actually occured. The similarity of item pairs is then
estimated each time the number of common users $n$ is incremented. In this way,
we can quantitatively compare various estimators by calculating their mean
squared error. Since different estimators may perform better for different
degrees of sparsity, we compute the mean squared error for $n=1,...,N$
separately.

@@

In our first approach, we model the probability distributions of true and
observed similarity. The goal is to use the observed similarity score directly
to approximate the true similarity. In this way, we can consider an abstact
$TrueSim$ which ignores the details of $TrueSimRand$ or $TrueSimPref$.
Similarly, we consider $Sim_n$ to be a generic observed similarity score when
the two items have $n$ common reviewers. Let $Y$ be a random variable
representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
random variable representing the $Sim_{n}$ of a pair of items when $n$ common
users are observed.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear predictors. For each number of
reviewers in common, $n$, the graphs represent the average squared error of
predicted $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the
technique of guessing the average $GoldSim$ is included. For these experiments,
we use $Sim$ = $PearSim$ and $N \geq 80$ for determining $GoldSim$.}
    \label{fig:Linear}
\end{figure}

@@

Let $Y$ be a random variable representing the $TrueSimRand$ of a randomly chosen
pair of items. As before, we model $Y$ as a Normal distribution
\begin{align}
Y \sim N(\mu_s, \sigma_s^2),
\end{align}
where $\mu_s$ and $\sigma_s^2$ are the average and variance of all true
similarity scores, respectively. Let $\vec{R} = \left(R_A~~R_B\right)$ be a
random two element vector representing the ratings given to items $A$ and $B$.
We model the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
distribution
\begin{align}
(\vec{R} | Y=y) \sim N(\vec{\mu_r}, \mathbf{\Sigma_y}),
\end{align}
where $\vec{\mu_r} = (\mu_r~~\mu_r)$ are the average ratings and
$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix.  Since
$y$ is the $TrueSimRand$ score, it can be thought of as the Pearson correlation
over the distribution of ratings $R_A | Y$ and $R_B | Y$. Assuming that the
variance of individual ratings $\sigma_r^2$ is independent of $y$, we have 
\begin{align}
y = \frac{\Cov{\left[R_A | Y=y , R_B | Y=y\right]}}{\sigma_r^2},
\end{align}
It therefore follows that if the ratings are indeed normally distributed 
then the covariance matrix is related to the true similarity $y$ by
\begin{align}
\mathbf{\Sigma_y} = \sigma_r^2 \cdot
\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right].
\end{align}

@@

As already mentioned, the estimator $PearSim$ is unable to say much about the
similarity of two items when only provided ratings from a single common user. In
this case, $PearSim$ yields $\pm1$ (or the indeterminate value $0/0$ in the case
that both ratings are zero) depending on whether the ratings have the same sign
or not. The value of $\hat{y}$ provided by the above model rectifies this by
generating similarity estimates in the range $[-1, +1]$ given a single pair of
ratings. We therefore consider each user who rates a pair of items as implicitly
predicting the similarity of the items. A key discovery of the present work is
that the accuracy of users' predictions in a training set is correlated with the
accuracy in a test set. We exploit this finding by ascribing a quantity we call
{\em user-predictivity} to each user defined as
\begin{align}\label{eqn:predictivity}
p_u = e^{-\epsilon_u/\lambda},
\end{align}
where $\lambda$ is a parameter to be learned and
\begin{align}\label{eqn:userError}
\epsilon_u = \frac{1}{\left|I_u\right|}
\sum_{A,B \in I_u}{\left|\hat{y}(r_{u,A},r_{u,B}) - GoldSim(A, B)\right|}
\end{align}
with $I_u$ the set of item pairs rated by user $u$. We are once again using
$GoldSim$ as a proxy for $TrueSim$ (See {\bf Experiment} section) such that
$\epsilon_u$ is the average absolute error of user $u$'s similarity predictions.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{RatingDist_40.png}
    \includegraphics[width=0.7\textwidth]{RatingDist_80.png}
	\caption{Distribution of rating pairs by common users for item pairs.  Top
figure shows distribution for item pairs having at least 40 common users. Bottom
figure shows distribution for item pairs having at least 80 common users.
Rating1 and Rating2 are chosen in arbitrary order.}
    \label{fig:RatingDist}
\end{figure}

@@

Having chosen values for $\mu_s$, $\sigma_s$, $\mu_r$, and $\sigma_r$ we are now
able to estimate the similarity of item pairs. To do this, we must solve Eq.
(\ref{eqn:estimator}) for $\hat{y}$ given a pair of ratings. This can be easily
done in practice using a root finding algorithm\footnote{To solve Eq.
(\ref{eqn:estimator}) for $\hat{y}$ given ratings $r_A$ and $r_B$, we use the
{\tt brentq} root finding function of the widely available Python module {\tt
scipy.optimize}.}. The implied predictions of individual users can then be
aggregated to form a new estimate of the similarity of items $A$ and $B$ as
\begin{align}\label{eqn:userPredSim}
UserPredSim(A,B) = \frac{\sum_{u \in U_{AB}}{p_u\cdot\hat{y}(r_{u,A}, r_{u,B})}}
{\sum_{u \in U_{AB}}{p_u}},
\end{align}
where, as before, $U_{AB}$ is the set of common users for items $A$ and $B$ and
$p_u$ is the {\em user-predictivity} defined in Eq. (\ref{eqn:predictivity}). We
leverage the training set to find $p_u$ for each user $u$. This is done by
recording the implied prediction $\hat{y}(r_{u,A},r_{u,B})$ for each common user
$u$ for each item pair $(A, B)$ in the training set and then computing each
user's average absolute error $\epsilon_u$ as per Eq. (\ref{eqn:userError}).
Finally, we apply supervised learning techniques to optimize the value of the
parameter $\lambda$ that appears in the definition of $p_u$. This is achieved by
minimizing the error in Eq. {\ref{eqn:userPredSim} made against the training
set. We find that $\lambda = 0.01$, which yields a strong preference for high
{\em user-predictivity}, works well for the data studied here.

Figure \ref{fig:MSE_SimR} shows the results of our experiment using
$UserPredSim$ to estimate $TrueSimRand$. To be clear, these plots are showing
the mean squared error (MSE) given by
\begin{align}
\mathrm{MSE} = \frac{1}{\left|S\right|}\sum_{(A,B) \in S}
{\left(Estimator(A,B) - GoldSim(A,B)\right)^2},
\end{align}
where $S$ is the test set of item pairs and $Estimator(A, B)$ represents an
arbitrary estimator of the similarity between items $A$ and $B$. In the upper
panel, we show the MSE for the $N \geq 40$ data. The black dashed line
represents the estimates of $PearSim$. For example, with $n = 20$ common users,
the MSE of $PearSim$ is about $0.07$, which corresponds to an average absolute
error of $0.26$ in the similarity score---roughly the magnitude of $\sigma_s$
(see Fig. \ref{fig:Histograms}). These data clearly show that the reliability of
$PearSim$ in estimating $TrueSimRand$ deteriorates rapidly as $n$ decreases.
Below about $n = 20$, the horizontal red line, which represents the MSE obtained
from always guessing the average $GoldSimRand$ value from the training set,
outperforms $PearSim$. These two estimators serve as simple benchmarks.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.9\textwidth]{MSE_SimR_40.png}
    \includegraphics[width=0.9\textwidth]{MSE_SimR_80.png}
	\caption{Mean square error {\em vs} number of common users, $n$, for
estimates of {\em TrueSimRand} for item pairs ultimately having at least 40
common users (top) and 80 common users (bottom). The dashed curve
represents {\em PearSim} computed using only the ratings from the first $n$
users. The flat red line indicate the MSE that results from always guessing
the average {\em TrueSimRand} from the training set. The green curve is {\em
LinRegSim} and the blue curve is {\em UserPredSim}. }
    \label{fig:MSE_SimR}
\end{figure}

@@

The lower panel of Fig. \ref{fig:MSE_SimR} shows the MSE of estimators in the
experiment carried out with item pairs with $N \geq 80$. The curves for
$PearSim$, always guessing the average, and $LinRegSim$ exhibit behavior
analogous to that observed in the $N \geq 40$ experiment, with addition of
increased noise. The performance of $UserPredSim$, meanwhile, is significantly
improved, achieving the MSE that $PearSim$ reaches at $n \approx 70$ by $n
\approx 20$. This result demonstrates the efficacy, at least in some datasets,
of measuring and exploiting {\em user-predictivity} when estimating item-item
similarity.

@@

The first is based on a model of noisy score observations. Using a series of
linear regressions, this estimator consistently performs as well or better than
both the Pearson correlation or cosine similarity at estimating future values of
these same functions. The second new estimator is based on a model of noisy
rating pairs generated by an item pair having a true, but unknown, similarity.
This method uses a parametrized surface to efficiently estimate item-item
similarity given only a single pair of ratings. Combining these implied
similarity ``predictions'' with the notion of {\em user-predictivity}, we
constructed a novel item-item similarity estimator that performs very well, but
whose effectiveness appears limited to estimations of future Pearson
correlation.

@@

In a different direction, we believe there are a number of possible applications
of the user-predictivity score presented in this paper. Based on the results
found here, it is reasonable to believe that adding weights to users by their
predictivity would increase the performance of any user-item rating predictor.
The predictivity could also have applications in the problem of Robust
Collaborative Filtering \cite{Mehta2007}, where rating data is particularly noisy
or riddled with spam.
